MATH181 Project: Automatic Identification of Nonlinear Systems

### 2024-04-30 Paper/main.tex
final (maybe?) paper

\section{Introduction} \label{sec:introduction}
Dynamical systems are mathematical objects used to describe anything and everything that changes. This gives rise to their utility in modeling a vast array of applications. Of course, physical systems are of particular interest as the rules of physics work in terms of velocity and acceleration---the descriptors of movement. 
Though, in principle, far more than the cannonical examples of physical movement are ``changing.'' Economic models describe the changes in producer and consumer supply and demand, while biologists model the rise and fall of organism populations in an ecosystem. 
The prevalence of these inherently changing systems leads to the field of dynamical systems having vast applicability to the world of mathematical modeling. Though while they are often the natural language to describe these situations, determining the particular dynamics in a given system is a fundamentally difficult problem.

This paper discusses a method through which data can be used to uncover the governing equations for a dynamical system. 
%
% This project discusses the procedure of recovering the underlying equations for the dynamics given noisy data collected by the system. So called ``physics-inspired'' or ``physics-informed'' AI.
% The goal of this field is to gain a physical understanding of systems where we can measure various trajectories, but the dynamics are unknown. Though this is additionally useful in situations where complex systems have dynamical properties that can be modeled on a smaller scale. 
In particular, the focus of this paper is on discussing the results demonstrated by Steven Brunton, Joshua Proctor, and Nathan Kutz in their paper published in 2016. They propose a new algorithm for reconstructing dynamics within a system called SINDy (Sparse Identification of Nonlinear Dynamics) \cite{sindy}.
% This first section aims to discuss the field more broadly and relevant papers to the results proposed in the aforementioned paper as well as more generally in the field. 
In this first section, I will discuss the field more broadly prior to Brunton, Proctor, and Kutz published their work. This will include a dicussion of relevant papers that provided the groundwork for the SINDy algorithm.
In the section section (\ref{sec:mathematical_background}), I will discuss the relevant mathematical background for SINDy and its underlying mathematics. Section \ref{sec:results} will go deeper into the results demonstrated by Brunton et al. by discussing my work to recreate the SINDy algorithm. 
Finally, section \ref{sec:discussion} offers a dicussion of the results in the paper and further areas of interest.

@@

\section{Mathematical Background} \label{sec:mathematical_background}
For the SINDy algorithm, the authors aimed to reframe the problem of recovering nonlinear dynamics using \textbf{sparse regression} and \textbf{compressed sensing} \cite{sindy}. Both sparse regression and compressed sensing broadly refer to the process of finding a regression solution to a problem such that the solution is sparse in some space \cite{sindy}. 
Brunton et al. relate that utilizing sparse regression is a strategy to reduce the noise present in the fitting of systems \cite{sindy}. The particular assumptions for sparse regression come from the idea that in most cases, the dynamics for a particular system will be governed by a small set of given basis functions. Therefore, the sparsity also gives advantages for computation as we can increase the size of the function space dramatically without needing significantly more computation \cite{sindy}.

% mention LASSO here as alternative method for sparse regression
The authors of \cite{sindy} propose an iterative algorithm for sparse regression. In particular, they start with a least-squares fit to the data, and then iteratively remove coefficients below a given threshold. By refitting with the remaining, nonzero terms iteratively, the algorithm can arrive at a sparse solution \cite{sindy}.

The SINDy algorithm relies on the availability of both sampled data, $\mathbf X$, and the sampled derivative, $\mathbf{\dot X}$. However, the authors note that the availability of derivative measurements are not guaranteed and therefore propose using the \textbf{total variation regularized derivative} to mitigate the noise created from numerically calculating $\mathbf {\dot X}$ from $\mathbf X$ \cite{sindy}. In particular, they cite the algorithm used by Rudin et al. \cite{rudin}. 
Here, Rudin et al. use a framework for computing a desired, denoised output, $u$, from a noisy input, $u_0$, by assuming that $u = u_0 + n$ for some white noise function $n$. It is assumed that $n$ has a mean of $0$ with some given standard deviation. 
The minmization problem corresponds to minimizing $\int_\Omega \sqrt{u_x^2 + u_y^2} \, \mathrm d A$, which is contrained by $\int_\Omega u \, \mathrm dA = \int_\Omega u_0 \, \mathrm dA$ under the assumption of the additive white noise. It is then proposed to solve for $u$ through either an iterative algorithm (i.e. simulated annealing) or, preferably, a PDE solver to find local minimum \cite{rudin}.

@@

\subsection{SINDy} \label{sec:SINDy}
The most basic systems considered for the algorithm SINDy are those of the form 
\[ \frac{\mathrm d}{\mathrm dt} \mathbf x(t) = \mathbf f (\mathbf x(t)). \]
In order to use the concepts of sparse regression and compressed sensing, discussed above, the authors assume that $\mathbf f$ is sparse in some function space \cite{sindy}.
The state vector, $\mathbf x(t)$, is collected over some time period and $\mathbf {\dot x}$ is either sampled alongside $\mathbf x(t)$ or numerically calculated from it. 
These samplings create the following matrices: 
\[ \mathbf X = \begin{bmatrix}
	\mathbf x^T(t_1) \\ \mathbf x^T(t_2) \\ \vdots \\ \mathbf x^T(t_m)
\end{bmatrix} = \begin{bmatrix}
	x_1(t_1) & x_2(t_1) & \cdots & x_n(t_1) \\
	x_1(t_2) & x_2(t_2) & \cdots & x_n(t_2) \\
	\vdots & \vdots & \ddots & \vdots \\ 
	x_1(t_m) & x_2(t_m) & \cdots & x_n(t_m)
\end{bmatrix}, \]  
and the immediately following $\mathbf {\dot X}$. We can then construct a library of candidate functions: 
\[  \Theta(\mathbf X) = \begin{bmatrix} \mid & \mid & \mid & \mid &  & \mid & \mid & \\ 1 & \mathbf X & {\mathbf X}^{P_2} & {\mathbf X}^{P_3} & \cdots & \sin{(\mathbf X)} & \cos{(\mathbf X)} & \cdots \\ \mid & \mid & \mid & \mid & & \mid & \mid & \end{bmatrix}.  \]
Here, $\mathbf X^{P_2}$ represents the quadratic nonlinearities in $\mathbf f$, in particular: 
\[ \mathbf X^{P_2} = \begin{bmatrix}
	x_1^2(t_1) & x_1(t_1)x_2(t_1) & \cdots & x_1(t_1)x_n(t_1) & x_2^2(t_1) & \cdots & x_n^2(t_1) \\
	x_1^2(t_2) & x_1(t_2)x_2(t_2) & \cdots & x_1(t_2)x_n(t_2) & x_2^2(t_2) & \cdots & x_n^2(t_2) \\
	\vdots & \vdots & \ddots & \vdots & \vdots & \ddots & \vdots \\
	x_1^2(t_m) & x_1(t_m)x_2(t_m) & \cdots & x_1(t_m)x_n(t_m) & x_2^2(t_m) & \cdots & x_n^2(t_m) \\
\end{bmatrix}. \]
Similarly, $\sin{(\mathbf X)}$ would be 
\[ \sin{(\mathbf X)} = \begin{bmatrix}
	\sin(x_1(t_1)) & \sin(x_2(t_1)) & \cdots & \sin(x_n(t_1)) \\ 
	\sin(x_1(t_2)) & \sin(x_2(t_2)) & \cdots & \sin(x_n(t_2)) \\
	\vdots & \vdots & \ddots & \vdots \\ 
	\sin(x_1(t_m)) & \sin(x_2(t_m)) & \cdots & \sin(x_n(t_m))
\end{bmatrix}. \]
Therefore, each column of $\Theta(\mathbf X)$ represents a candidate function for $\mathbf f$. From this, we set up the sparse regression problem
\[ \mathbf {\dot X} = \Theta(\mathbf X)\mathbf \Xi. \]
In particular, $\mathbf \Xi = \begin{bmatrix}
	\boldsymbol \xi_1 & \boldsymbol \xi_2 & \cdots & \boldsymbol \xi_n
\end{bmatrix}$ where each $\boldsymbol \xi_i$ is a sparse vector dictating which functions in the function library are active for each of the row equations in $\mathbf f$. The sparse regression problem is an iterative algorithm that iteratively thresholds out coefficients in $\mathbf \Xi$ under some sparsity parameter, $\lambda$, and then refits the problem on the remaining functions.

The authors additionally suggest a slight modification to the algorithm in order to detect bifurcations present in a system. In particular, the authors appended the bifurcation parameter to the dynamics: 
\begin{equation} \label{eq:sindy_with_bifurcation}
	\begin{split}
		\mathbf {\dot x} &=  \mathbf{f}(\mathbf x; \mu), \\
		\dot \mu &= 0.
	\end{split}
\end{equation}
Then, by sampling data from different values of $\mu$, it is possible for SINDy to reconstruct the system from combinations of $\mathbf x$ as well as $\mu$. In turn, the system can then find the dynamics present with changes in $\mu$, and hence bifurcations.

\section{Results} \label{sec:results}
\subsection{Code}
The main object in recreating the results presenting in the paper was writing a codebase that is able to recreate the SINDy algorithm. 
I wrote the code in the Julia Programming Language \cite{julia} and with all plotting done in PlotlyBase.jl \cite{PlotlyBase.jl}.

All data used to train the SINDy model was collected by integrating using ordinary differential equation (ODE) solvers provided by the DifferentialEquations.jl \cite{rackauckas2017differentialequations} package.
For the numerical stability of the model, data was collected at fixed timesteps, as shown in Listing \ref{lorenz_data} for the Lorenz system.

\jlinputlisting[language = Julia, caption = {Lorenz system data simulation}, label = {lorenz_data}, firstline = 9, lastline = 14]{../Code/DynamicsRecovery/src/Systems/lorenz.jl}

Additionally, noise sampled from a standard normal distribution was scaled by a noise parameter and added to the collected data before SINDy was trained on it. This code utilized the Distributions.jl \cite{JSSv098i16}\cite{Distributions.jl-2019} package to provide the normal distribution. 
Once the data was collected and noise applied, the program constructed the data matrix, $\mathbf X$, and the corresponding derivative matrix, $\mathbf {\dot X}$. Differentiation is done in one of two ways, either through direct numerical differentiation, using Interpolations.jl \cite{Interpolations.jl}, or using the total-variation regularization derivative, provided by NoiseRobustDifferentiation.jl \cite{chartrand2011numerical}\cite{NoiseRobustDifferentiation.jl}.
Using a linear interpolation to compute the derivatives had the additional benefit of being more stable when samples were not taken with constant timestamps, and deviates in approach from the code written by Brunton et al. \cite{sindy}. 
The code for constructing the data and derivative libraries is shown in Listing \ref{X_Xdot_construction}.

% https://tex.stackexchange.com/questions/73231/avoid-page-breaks-in-lstlistings
% https://tex.stackexchange.com/questions/39017/how-to-influence-the-position-of-float-environments-like-figure-and-table-in-lat
\jlinputlisting[language = Julia, caption = {Data and derivative library construction}, label = {X_Xdot_construction}, firstline = 68, lastline = 95, float, floatplacement = tb]{../Code/DynamicsRecovery/src/SINDy/data_ingestion.jl}

@@

Finally, the sparse regression algorithm to compute $\mathbf \Xi$ is the same process as outlined by Brunton et al. \cite{sindy}. In particular, the starting point for the iterative algorithm is a least-squares solution to the problem using a standard regression. Then, over $k$ iterations, coefficients in $\mathbf \Xi$ which are less than the sparsity parameter, $\lambda$, are dropped and replaced with zero. In each of these iterations, the regression is then recomputed with only the nonzero entries. 
This code is presented in Listing \ref{sparse_regression}.

\jlinputlisting[language = Julia, caption = {Sparse regression}, label = {sparse_regression}, firstline = 3, lastline = 23, float, floatplacement = ptb]{../Code/DynamicsRecovery/src/SINDy/sparse_regression.jl}

The resulting system defined by $\mathbf \Xi$ can then be solved using the same methods as when collecting data for the original system. The governing equations are constructed using the same function generation from the library construction in order to ensure complete generality.

@@

	\centering
	\includegraphics[width = 0.47\textwidth, trim={4.3cm, 1.9cm, 4.3cm, 3.5cm}, clip]{lorenz_actual_graph.png}
\end{figure}

I ran the SINDy algorithm on both data with noise of magnitude $\eta = 0.01$ and $\eta = 0.1$. Of interest is the system with $\eta = 0.1$ was only able to converge on a solution when the derivative was computed using the total-variation regularization derivative. 
For a function basis, I chose a basis comprised of all 7th-degree polynomial terms and all sines and cosines of the form $\sin{(ku_i)}$ and $\cos{(ku_i)}$ for all $\{u_i\} = \{x,y,z\}$ and integer $k$ from $1$ to $10$. This created a basis of $180$ possible functions. 
Further, the sparsification parameter was chosen to be $\lambda = 0.25$ with 10 iterations for sparse regression.
Integration of the input data was done over a timespan from $0$ to $100$ with $dt = 0.001$. 
The resulting systems from $\eta = 0.01$ are shown in Figure \ref{fig:lorenz_system_found_001} and the system resulting from $\eta = 0.1$ is shown in Figure \ref{fig:lorenz_system_found_01}.

\begin{figure}[h]
	\caption{Discovered Lorenz system with noise $\eta = 0.01$}
	\label{fig:lorenz_system_found_001}

	\centering 
	\begin{minipage}{0.47\textwidth}
		\begin{center}
			\includegraphics[width = \textwidth, trim={4.3cm, 1.9cm, 4.3cm, 3.5cm}, clip]{lorenz_found_001_normal.png}

			With standard derivative calculation
		\end{center}
	\end{minipage}%
	\begin{minipage}{0.47\textwidth}
		\begin{center}
			\includegraphics[width = \textwidth, trim={4.3cm, 1.9cm, 4.3cm, 3.5cm}, clip]{lorenz_found_001_tvdiff.png}

			With total-variation regularization derivative
		\end{center}
	\end{minipage}
\end{figure}

\begin{figure}[h]
	\caption{Discovered Lorenz system with noise $\eta = 0.1$ and total-variation regularization differentiation}
	\label{fig:lorenz_system_found_01}

	\centering
	\includegraphics[width = 0.47\textwidth, trim={4.3cm, 1.9cm, 4.3cm, 3.5cm}, clip]{lorenz_found_01_tvdiff.png}
\end{figure}

>>>>

\section{Introduction} \label{sec:introduction}
Dynamical systems are mathematical objects used to describe anything and everything that changes. This gives rise to their utility in modeling a vast array of real-world systems. Of course, physical systems are of particular interest as the rules of physics work in terms of velocity and acceleration---the descriptors of movement. 
Though, in principle, far more than the cannonical examples of physical movement are ``changing.'' Economic models describe the changes in producer and consumer supply and demand, while biologists model the rise and fall of organism populations in an ecosystem. 
The prevalence of these inherently changing systems leads to the field of dynamical systems having vast applicability to the world of mathematical modeling. Though while they are often the natural language to describe these situations, determining the particular dynamics in a given system is a fundamentally difficult problem.

This paper discusses a method through which data can be used to uncover the governing equations for a dynamical system. 
%
% This project discusses the procedure of recovering the underlying equations for the dynamics given noisy data collected by the system. So called ``physics-inspired'' or ``physics-informed'' AI.
% The goal of this field is to gain a physical understanding of systems where we can measure various trajectories, but the dynamics are unknown. Though this is additionally useful in situations where complex systems have dynamical properties that can be modeled on a smaller scale. 
In particular, the focus of this paper is on discussing the results demonstrated by Steven Brunton, Joshua Proctor, and Nathan Kutz in their paper published in 2016. They propose a new algorithm for reconstructing dynamics within a system called SINDy (Sparse Identification of Nonlinear Dynamics) \cite{sindy}.
% This first section aims to discuss the field more broadly and relevant papers to the results proposed in the aforementioned paper as well as more generally in the field. 
In this first section, I will discuss the field more broadly prior to Brunton, Proctor, and Kutz publishing their work. This will include a dicussion of relevant papers that provided the groundwork for the SINDy algorithm.
I will then go on to describe relevant mathematical background for SINDy and its underlying mathematics (Section \ref{sec:mathematical_background}). Section \ref{sec:results} will go deeper into the results demonstrated by Brunton et al. by discussing my work to recreate the SINDy algorithm as well as further exploration into the performance of SINDy. 
Finally, section \ref{sec:discussion} offers a dicussion of the results in the paper and further areas of interest.

@@

\section{Mathematical Background} \label{sec:mathematical_background}
For the SINDy algorithm, the authors aimed to reframe the problem of recovering nonlinear dynamics using \textbf{sparse regression} and \textbf{compressed sensing} \cite{sindy}. Both sparse regression and compressed sensing broadly refer to the process of finding a regression solution to a problem such that the solution is sparse in some function space \cite{sindy}. 
Brunton et al. relate that utilizing sparse regression is a strategy to reduce the noise present in the fitting of systems \cite{sindy}. The particular assumptions for sparse regression come from the idea that in most cases, the dynamics for a particular system will be governed by a small set of given basis functions. Therefore, the sparsity also gives advantages for computation as we can increase the size of the function space dramatically without needing significantly more computation \cite{sindy}.

% mention LASSO here as alternative method for sparse regression
% The authors of \cite{sindy} propose an iterative algorithm for sparse regression. In particular, they start with a least-squares fit to the data, and then iteratively remove coefficients below a given threshold. By refitting with the remaining, nonzero terms iteratively, the algorithm can arrive at a sparse solution \cite{sindy}.

The SINDy algorithm relies on the availability of both sampled data, $\mathbf X$, and the sampled derivative, $\mathbf{\dot X}$. However, the authors note that the availability of derivative measurements are not guaranteed and therefore propose using the \textbf{total-variation regularized derivative} to mitigate the noise created from numerically calculating $\mathbf {\dot X}$ from $\mathbf X$ \cite{sindy}. 
In particular, the total-variation regularization technique is attributed to Rudin et al. \cite{rudin}. This method for computing derivatives is robust to noise present in a system, without affect the performance in identifying jump discontinuities \cite{chartrand2011numerical}. 
The implementation of the total-variation regularization derivative is beyond the scope of the paper. 
% Here, Rudin et al. use a framework for computing a desired, denoised output, $u$, from a noisy input, $u_0$, by assuming that $u = u_0 + n$ for some white noise function $n$. It is assumed that $n$ has a mean of $0$ with some given standard deviation. 
% The minmization problem corresponds to minimizing $\int_\Omega \sqrt{u_x^2 + u_y^2} \, \mathrm d A$, which is contrained by $\int_\Omega u \, \mathrm dA = \int_\Omega u_0 \, \mathrm dA$ under the assumption of the additive white noise. It is then proposed to solve for $u$ through either an iterative algorithm (i.e. simulated annealing) or, preferably, a PDE solver to find local minimum \cite{rudin}.

@@

\subsection{SINDy} \label{sec:SINDy}
The main purpose of SINDy is in identifying the function $\mathbf f$ in (\ref{eq:dynamical_system}). 
\begin{equation}\label{eq:dynamical_system}
	\frac{\mathrm d}{\mathrm dt} \mathbf x(t) = \mathbf f (\mathbf x(t)).
\end{equation}
Such identification requires the collection of timeseries data, $\mathbf X$. The form of $\mathbf X$ is as follows:

% In order to use the concepts of sparse regression and compressed sensing, discussed above, the authors assume that $\mathbf f$ is sparse in some function space \cite{sindy}.
% The state vector, $\mathbf x(t)$, is collected over some time period and $\mathbf {\dot x}$ is either sampled alongside $\mathbf x(t)$ or numerically calculated from it. 
% These samplings create the following matrices: 
\[ \mathbf X = \begin{bmatrix}
	\mathbf x^T(t_1) \\ \mathbf x^T(t_2) \\ \vdots \\ \mathbf x^T(t_m)
\end{bmatrix} = \begin{bmatrix}
	x_1(t_1) & x_2(t_1) & \cdots & x_n(t_1) \\
	x_1(t_2) & x_2(t_2) & \cdots & x_n(t_2) \\
	\vdots & \vdots & \ddots & \vdots \\ 
	x_1(t_m) & x_2(t_m) & \cdots & x_n(t_m)
\end{bmatrix}, \]  
where each row of $\mathbf X$ is some measurement in time of the $n$ state variables in our system. Hence, the columns of $\mathbf X$ correspond to the different state variables of our system, and the rows of $\mathbf X$ represent timeseries data measurements from our system. Either sampled alongside $\mathbf X$, or numerically computed from it, we also need the relevant derivatives, 
\[ \mathbf {\dot X} = \begin{bmatrix}
	\mathbf {\dot x}^T(t_1) \\ \mathbf {\dot x}^T(t_2) \\ \vdots \\ \mathbf {\dot x}^T(t_m)
\end{bmatrix} = \begin{bmatrix}
	\dot x_1(t_1) & \dot x_2(t_1) & \cdots & \dot x_n(t_1) \\
	\dot x_1(t_2) & \dot x_2(t_2) & \cdots & \dot x_n(t_2) \\
	\vdots & \vdots & \ddots & \vdots \\ 
	\dot x_1(t_m) & \dot x_2(t_m) & \cdots & \dot x_n(t_m)
\end{bmatrix}. \]  
As discussed above, Brunton et al. propose using the total-variation regularization derivative to combat noise if numerically calculating $\mathbf {\dot X}$ from $\mathbf X$. 
It is important to note that no assumptions are made about the time steps between each $t_i$, nor do all collected data points need to be from the same orbit.

In order to specify the particular function basis we aim to use for SINDy, we construct a function library. In essence, this library is constructed such that each column is representative of a candidate function in our function basis, 
\[ \Theta(\mathbf X) = \begin{bmatrix}
	g_1(\mathbf x(t_1)) & g_2(\mathbf x(t_1)) & \cdots & g_k(\mathbf x(t_1)) \\ 
	g_1(\mathbf x(t_2)) & g_2(\mathbf x(t_2)) & \cdots & g_k(\mathbf x(t_2)) \\ 
	\vdots & \vdots & \ddots & \vdots \\ 
	g_1(\mathbf x(t_m)) & g_2(\mathbf x(t_m)) & \cdots & g_k(\mathbf x(t_m)) 
\end{bmatrix}. \]
In principle, each $g_i$ could be any function in terms of any subset of the state variables. Computation of $\Theta(\mathbf X)$ simply amounts to the evaluation of these candidate functions on all timeseries measurements from our system. 
We will seek a solution for the governing equation of each state variable as some linear combination of these candidate functions, 
\[ \dot x_i(t) = \xi_{i, 1}g_1(\mathbf x(t)) + \xi_{i, 2}g_2(\mathbf x(t)) + \cdots + \xi_{i, k}g_k(\mathbf x(t)). \]
In particular, we seek a sparse solution, meaning that most of the coefficients are zero (most $\xi_{i, j} = 0$). 
In practice, our function library could contain a mixture of polynomial terms and sines or cosines,

\[  \Theta(\mathbf X) = \begin{bmatrix} \mid & \mid & \mid & \mid &  & \mid & \mid & \\ 1 & \mathbf X & {\mathbf X}^{P_2} & {\mathbf X}^{P_3} & \cdots & \sin{(\mathbf X)} & \cos{(\mathbf X)} & \cdots \\ \mid & \mid & \mid & \mid & & \mid & \mid & \end{bmatrix}.  \]
Here, $\mathbf X^{P_2}$ represents the quadratic nonlinearities in $\mathbf f$, 
\[ \mathbf X^{P_2} = \begin{bmatrix}
	x_1^2(t_1) & x_1(t_1)x_2(t_1) & \cdots & x_1(t_1)x_n(t_1) & x_2^2(t_1) & \cdots & x_n^2(t_1) \\
	x_1^2(t_2) & x_1(t_2)x_2(t_2) & \cdots & x_1(t_2)x_n(t_2) & x_2^2(t_2) & \cdots & x_n^2(t_2) \\
	\vdots & \vdots & \ddots & \vdots & \vdots & \ddots & \vdots \\
	x_1^2(t_m) & x_1(t_m)x_2(t_m) & \cdots & x_1(t_m)x_n(t_m) & x_2^2(t_m) & \cdots & x_n^2(t_m) \\
\end{bmatrix}, \]
while $\sin{(\mathbf X)}$ would be 
\[ \sin{(\mathbf X)} = \begin{bmatrix}
	\sin(x_1(t_1)) & \sin(x_2(t_1)) & \cdots & \sin(x_n(t_1)) \\ 
	\sin(x_1(t_2)) & \sin(x_2(t_2)) & \cdots & \sin(x_n(t_2)) \\
	\vdots & \vdots & \ddots & \vdots \\ 
	\sin(x_1(t_m)) & \sin(x_2(t_m)) & \cdots & \sin(x_n(t_m))
\end{bmatrix}. \]
After constructing our function library, we may continue by solving the sparse regression problem, 
\[ \mathbf {\dot X} = \Theta(\mathbf X)\mathbf \Xi. \]
In particular, we aim to find a $\mathbf \Xi$, similar to a traditional least-squares regression, such that $\mathbf \Xi$ is sparse. 
A further discussion of the mechanisms for performing this sparse regression are discussed in Section \ref{sec:code}. 
We note that $\mathbf \Xi = \begin{bmatrix}
	\boldsymbol \xi_1 & \boldsymbol \xi_2 & \cdots & \boldsymbol \xi_n
\end{bmatrix}$ where each $\boldsymbol \xi_i$ is a sparse vector dictating the coefficients on the candidate functions for the governing equation of state variable $x_i$.

The authors additionally suggest an alteration to the algorithm in order to detect bifurcations present in a system. In particular, the authors appended the bifurcation parameter to the dynamics: 
\begin{equation} \label{eq:sindy_with_bifurcation}
	\begin{split}
		\mathbf {\dot x} &=  \mathbf{f}(\mathbf x; \mu), \\
		\dot \mu &= 0.
	\end{split}
\end{equation}
Then, by sampling data from different values of $\mu$, it is possible for SINDy to reconstruct the system from combinations of $\mathbf x$ as well as $\mu$. 
This results in an identified system that can capture the bifurcation behavior in changes of $\mu$.

\section{Results} \label{sec:results}
\subsection{Code} \label{sec:code}
Central to recreating the results presented by Brunton et al. was writing a novel implementation of the SINDy algorithm.
I wrote the code in the Julia Programming Language \cite{julia}.
All data used to train the SINDy model was collected by integrating using ordinary differential equation (ODE) solvers provided by the DifferentialEquations.jl \cite{rackauckas2017differentialequations} package.

In order to replicate sensor noise, randomly sampled noise from a standard normal distribution was added to all collected data ($\mathbf X$).
This noise was scaled by some noise mangitude, $\eta$, in order to test the performance of SINDy at various levels of system noise. Additionally, all derivatives $(\mathbf {\dot X}$) were numerically calculated from the data after noise was applied. 
Distributions.jl \cite{JSSv098i16}\cite{Distributions.jl-2019} was used to sample the noise and NoiseRobustDifferentiation.jl \cite{chartrand2011numerical}\cite{NoiseRobustDifferentiation.jl} provided an implementation of the total-variation regularization derivative in order to compute $\mathbf{\dot X}$.

@@

Finally, the sparse regression algorithm to compute $\mathbf \Xi$ is the same process as outlined by Brunton et al. \cite{sindy}. The code for this algorithm is shown in Listing \ref{lst:sparse_regression}.
In particular, the algorithm starts by taking an initial ordinary least squares solution for $\mathbf \Xi$. Following this, the algorithm iteratively finds and removes coefficients in $\mathbf \Xi$ under some sparsity parameter, $\lambda$. On each iteration, after finding these small coefficients, we set these coefficients to zero and re-regress on only the remaining non-zero terms. 
Brunton et al. found that this algorithm converges rapidly (for a few iteration steps), meaning that 5-10 iterations are often sufficient.

% \jlinputlisting[language = Julia, caption = {Sparse regression}, label = {sparse_regression}, firstline = 3, lastline = 23, float, floatplacement = ptb]{../Code/DynamicsRecovery/src/SINDy/sparse_regression.jl}
\begin{jllisting}[language = Julia, caption = {Sparse regression}, label = {lst:sparse_regression}, float, floatplacement = ptb]
function sparse_regression(Θ::Matrix{Float64}, Xdot::Matrix{Float64}, λ::Float64, kmax::Integer) 
    # create initial guess for Xi as just regression
    Ξ = Θ\Xdot

    for _ in 1:kmax 
        # find coefficients that are less than λ 
        small_coefficients = abs.(Ξ) .< λ

        # set small coefficients to 0 
        Ξ[small_coefficients] .= 0

        # now individually run regression on non-zero coefficients
        for n in 1:(size(Xdot)[2]) # loop through all state dimensions 
            nonsmall = .!small_coefficients[:, n]

            Ξ[nonsmall, n] = Θ[:, nonsmall]\Xdot[:, n]
        end # for n
    end # for k

    Ξ
end # function sparse_regression
\end{jllisting}
After this process, $\mathbf \Xi$ defines a ``discovered system'' from the data. In particular, we can use this discovered system to project dynamics for the original system.

% The resulting system defined by $\mathbf \Xi$ can then be solved using the same methods as when collecting data for the original system. The governing equations are constructed using the same function generation from the library construction in order to ensure complete generality.

\subsection{Measuring Error} \label{sec:measuring_error}
In order to discuss the performance of SINDy, I aim to propose an error metric that accounts for the overall scale of the system, as well as a desire to improve the accuracy of the derivative over that of the orbit. A particular case study for the importance of emphasizing derivative accuracy over orbit accuracy is that of a chaotic system, where it would be reasonable to expect large deviations in orbits with derived systems closely matching the actual, underlying system. 
Therefore, qualitatively similar behavior could be labeled as inaccurate due to expected deviations from small variations in parameters. Further, we want an error metric to be proportional to the overall scale of the system, as scaling the coordinate system shouldn't change the accuracy of the derived system. 
For fulfilling these requirements, the following error metric may be used: 
\[ \varepsilon = \sqrt{ \left( \sum_{i \in T} \frac{(\dot x_1(t_i) - f_1(t_i))^2}{\sum_{i \in T} (\dot x_1(t_i))^2} \right)^2 + \cdots + \left( \sum_{i \in T} \frac{(\dot x_n(t_i) - f_n(t_i))^2}{\sum_{i \in T} (\dot x_n(t_i))^2} \right)^2 }, \]
where $T$ is a set of indices for the collected timeseries data, $\dot x_j(t_i)$ is the measured derivative from the system for state variable $j$ and time $i$, $f_j(t_i)$ is the same but for the found sytem, and $n$ is the state dimension. 
Typically, cross-validation techniques are used to test the accuracy of a fit, where a certain proportion of the collected data is withheld in order to calculate this error. Hence, $T$ represents that subset of data withheld for testing. 
In particular, this error metric can be thought of as how large, in proportion to the magnitude of the derivative of the system, the error in the derivative is for the found system. Throughout this paper I will cite this error metric for found systems to give an idea of the relative performance of SINDy between cases. 
As a standard, error will be calculated on 10 percent of the data randomly selected to be withheld for testing.

@@

	\centering
	\includegraphics[width = 0.47\textwidth, trim={4.3cm, 1.6cm, 4.3cm, 2.5cm}, clip]{lorenz_attractor.pdf}
\end{figure}

In this section, I will examine the performance of the SINDy algorithm across different magnitudes of noise applied to sampled points from the system. 
In particular, I will demonstrate results for noise magnitudes $\eta = 0.01$, $\eta = 0.1$, and $\eta = 1$ where noise is generated by sampling from a standard normal distribution, and scaling the sampled noise by $\eta$. 
Further, it is important to note that noise is applied to the data matrix, $\mathbf X$, before the derivatives, $\mathbf {\dot X}$ is computed. 
The derivatives are computed numerically from the sampled data after noise application, so in order to counteract the noise in the data, derivatives are calculated using the total-variation regularization derivative, as suggested by Brunton et al. \cite{sindy}. 
Data was collected with a timespan from $0$ to $100$, with $dt = 0.001$. For measuring error, $10\%$ of the collected data was withheld as testing data to compare against. 
The results for the different levels of noise can be seen in Figure \ref{fig:found_lorenz_systems}.
\begin{figure}[h]
	\caption{Discovered Lorenz systems with noise magnitudes $\eta = 0.01$, $\eta = 0.1$, and $\eta = 1$. Error ($\varepsilon$) is reported as described in Section \ref{sec:measuring_error}.}
	\label{fig:found_lorenz_systems}

	\centering 
	\begin{minipage}{0.32\textwidth}
		\[ \eta = 0.01, \, \lambda = 0.25, \]
		\[ \varepsilon = 0.00041 \]
		\begin{center}
			\includegraphics[width = 0.95\textwidth, trim={4.3cm, 1.6cm, 4.3cm, 2.5cm}, clip]{found_lorenz_001.pdf}
		\end{center}
		\resizebox{!}{1.75em}{$\begin{aligned} 
			\dot x &=-9.945x + 9.975y \\ 
			\dot y &= 27.063x  -0.674y  -0.977xz\\ 
			\dot z &= -2.649z + 0.996xy
		\end{aligned}$}
	\end{minipage}%
	\begin{minipage}{0.32\textwidth}
		\[ \eta = 0.1, \, \lambda = 0.4, \]
		\[ \varepsilon = 0.00301 \]
		\begin{center}
			\includegraphics[width = 0.95\textwidth, trim={4.3cm, 1.6cm, 4.3cm, 2.5cm}, clip]{found_lorenz_01.pdf}
		\end{center}
		\resizebox{!}{1.75em}{$\begin{aligned} 
			\dot x &= -9.800x + 9.832y \\ 
			\dot y &= 25.948x  -0.438y -0.947xz \\ 
			\dot z &= -2.615z + 0.983xy
		\end{aligned} $}
	\end{minipage}%
	\begin{minipage}{0.32\textwidth}
		\[ \eta = 1, \, \lambda = 0.416, \]
		\[ \varepsilon = 0.21021 \]
		\begin{center}
			\includegraphics[width = 0.95\textwidth, trim={4.3cm, 1.6cm, 4.3cm, 2.5cm}, clip]{found_lorenz_1.pdf}
		\end{center}
		\resizebox{!}{1.75em}{$\begin{aligned} 
			\dot x &= -7.007x + 7.148y  \\ 
			\dot y &= 14.836x + 1.578y  -0.631xz \\ 
			\dot z &=   -2.149z + 0.839xy -2.150
		\end{aligned}$}
	\end{minipage}
\end{figure}

We can see that for the three noise levels, SINDy was able to correctly identify the qualitative behavior of the Lorenz system. 
Further, SINDy identifies exclusively the correct terms in the systems for $\eta = 0.01$ and $\eta = 0.1$, deviating from the expected terms in $\eta = 1$ by a forcing term in the $\dot z$ equation (Figure \ref{fig:found_lorenz_systems}). 
Additionally, we notice that error in the produced system increases with noise, as we may expect, but additionally that this increase in error is not linear with noise. In particular, while the $\eta = 0.1$ system has approximately 10-fold the error as the $\eta = 0.01$ system, the $\eta = 1$ system sees an 100-fold increase in error over $\eta = 0.1$.

For completeness, we note that for the $\eta = 0.01$ system, we use the basis $$\mathcal B = \{ \mathbf 1, \mathbf X, \mathbf X^{P_2}, \ldots, \mathbf X^{P_7}, \sin(1\mathbf X), \ldots, \sin{10\mathbf X}, \cos{1\mathbf X}, \ldots, \cos{10\mathbf X} \}.$$ We use a smaller basis for $\eta = 0.1$, $$\mathcal B = \{  \mathbf 1, \mathbf X, \mathbf X^{P_2}, \ldots, \mathbf X^{P_5}, \sin(1\mathbf X), \ldots, \sin{10\mathbf X}, \cos{1\mathbf X}, \ldots, \cos{10\mathbf X}\},$$ 
and a smaller basis still for $\eta = 1$, 
\[ \mathcal B = \{ \mathbf 1, \mathbf X, \mathbf X^{P_2}, \ldots, \mathbf X^{P_5} \}. \]
The choice of basis was motivated by accuracy concerns with the higher levels of noise. For a larger discussion of the basis choice, see Appendix \ref{apx:lorenz_with_diff_bases}.
